Classical density functional theory for the square-well liquid

### 2015-03-04 papers/square-well-fluid/paper.tex
rename papers/square-well-liquid to papers/histogram and fix fac up Tested on darcy. Took 15:24.96.
Added another Fourier transform, and clarified some math in square well fluid paper. Tested on knightley. Took 0:30.22.

\newcommand\kk{\mathbf{k}}
\newcommand\xx{\mathbf{x}}
\newcommand\rr{\mathbf{r}}
\newcommand\fixme[1]{\textcolor{red}{\textbf{[#1]}}}

@@

A different way to do this that avoids some of the Fourier transform
troubles is to formulate $g_S$ as a polynomial expansion.  We do this
up to fouth order so that
\begin{align}
  g_S(r;g_\sigma) &=
  g_\sigma + \sum_{i=1}^{4} \sum_{j=1}^{4} \kappa_{ij} (g_\sigma - 1)^i
  \left(\frac{r}{\sigma}-1\right)^j,
  \label{eq:fit-form}
\end{align}
which makes the transforms for $i=1,2,3,4$
\begin{align}
 \frac {4\pi}{k}\int_\sigma^{\lambda \sigma} r
 \left( \frac{r}{\sigma}-1\right)^i \sin (kr) \mathrm{d}r.
\end{align}
Let's do a u-substitution where $u= \frac{r}{\sigma}-1$, then the
integral becomes
\begin{align}
 \frac {4\pi}{k}\int_0^{\lambda -1} \sigma^2 (u+1)u^i \sin(k\sigma
       [u+1]) \mathrm{d}u.
\end{align}
Now we calculate the integrals.
\begin{widetext}
\begin{align}
  \int_\sigma^{\lambda \sigma} r
 \left( \frac{r}{\sigma}-1\right) \sin (kr) \mathrm{d}r =&
 \frac{1}{k^3 \sigma}\left[(2-\lambda k^2
 \sigma^2(\lambda-1))\cos(k\lambda\sigma)-k\sigma(\sin(k\sigma)
 +(1-2\lambda) \sin(k\lambda\sigma)) -\cos(k\sigma) \right] \\
  \int_\sigma^{\lambda \sigma} r
 \left( \frac{r}{\sigma}-1\right)^2 \sin (kr) \mathrm{d}r =&
 \frac{1}{k^4\sigma^2} \big[
 6\sin(k\sigma)+(k^2\sigma^2(1-4\lambda+3\lambda^2)
 -6)\sin(k\lambda\sigma) \notag \\
 &-2k\sigma\cos(k\sigma) - k\sigma(4 +
 \lambda(k^2\sigma^2(\lambda-1)^2 - 6))\cos(k\lambda\sigma)\big] \\
  \int_\sigma^{\lambda \sigma} r
 \left( \frac{r}{\sigma}-1\right)^3 \sin (kr) \mathrm{d}r =&
 \frac{1}{k^5\sigma^3} \big[k\sigma(6\sin(k\sigma) + (18-24\lambda
   +k^2\sigma^2(\lambda-1)^2(4\lambda-1))\sin(k\lambda\sigma)) \notag \\
   &+24\cos(k\sigma) + (6k^2\sigma^2(\lambda-1)(2\lambda-1)-\lambda
   k^4\sigma^4(\lambda-1)^3)\cos(k\lambda\sigma)\big] \\
  \int_\sigma^{\lambda \sigma} r
 \left( \frac{r}{\sigma}-1\right)^4 \sin (kr) \mathrm{d}r =&
 \frac{1}{k^6\sigma^4}
 \big[(k\sigma(\lambda-1)(36-60\lambda+k^2\sigma^2(\lambda-1)^2(5\lambda-1))
   -1)\sin(k\lambda\sigma) -120\sin(k\sigma) \notag \\ 
   &+k\sigma(24\cos(k\sigma) -
   (24(5\lambda-4)-4k^2\sigma^2(\lambda-1)^2 (5\lambda-2) +
   \lambda k^4\sigma^4(\lambda-1)^4)\cos(k\lambda\sigma) \big]
\end{align}
\end{widetext}
\subsubsection{Averaged pair distribution  $g_\sigma(\rr)$}

@@

\fixme{We have introduced an approximation to the pair distribution
  function above... here we discuss an implementation of thermodynamic
  perturbation theory for an inhomogeneous system using this pair
  distribution function.}

>>>>

\newcommand\kk{\mathbf{k}}
\newcommand\xx{\mathbf{x}}
\newcommand\rr{\mathbf{r}}
\newcommand\dd{\mathrm{d}}
\newcommand\fixme[1]{\textcolor{red}{\textbf{[#1]}}}

@@

A different way to do this that avoids some of the Fourier transform
troubles is to formulate $g_S$ as a polynomial expansion.  We do this
up to fouth order so that
\begin{align}
  g_S(r;g_\sigma) &= g_\sigma + \sum_i \gamma_i(g_\sigma) \xi_i(r) \\
  &=
  g_\sigma + \sum_{i=1}^{4} \left(\sum_{j=1}^{4} (g_\sigma
  - 1)^j \kappa_{ji} \right)
  \left(\frac{r}{\sigma}-1\right)^i,
  \label{eq:fit-form}
\end{align}
which makes the transforms for $i=1,2,3,4$
\begin{align}
 \frac {4\pi}{k}\int r \xi_i(r) \Phi(r) \sin (kr) \dd r =  \frac
       {-4\pi\epsilon}{k}\int_\sigma^{\lambda \sigma} r \xi_i(r)\sin (kr) \dd r
\end{align}
For the ``zeroth'' term, we treat it like $\xi=1$, so that we're only
transforming the potential well.
%% For the other i's, we'll do a
%% u-substitution where $u= \frac{r}{\sigma}-1$, then the
%% integral becomes
%% \begin{align}
%%  \frac {4\pi}{k}\int_0^{\lambda -1} \sigma^2 (u+1)u^i \sin(k\sigma
%%        [u+1]) \mathrm{d}u.
%% \end{align}
%% Now we calculate the integrals.
\begin{widetext}
\begin{align}
  \int_\sigma^{\lambda \sigma} r \sin (kr) \mathrm{d}r =&
  \frac{1}{k^2}\left[ \sin (k\lambda\sigma) - k\lambda\sigma
    \cos (k\lambda\sigma) - \sin (k\sigma) + k\sigma \cos (k-sigma) \right]\\
  \int_\sigma^{\lambda \sigma} r
 \left( \frac{r}{\sigma}-1\right) \sin (kr) \mathrm{d}r =&
 \frac{1}{k^3 \sigma}\left[(2-\lambda k^2
 \sigma^2(\lambda-1))\cos(k\lambda\sigma)-k\sigma(\sin(k\sigma)
 +(1-2\lambda) \sin(k\lambda\sigma)) -\cos(k\sigma) \right] \\
  \int_\sigma^{\lambda \sigma} r
 \left( \frac{r}{\sigma}-1\right)^2 \sin (kr) \mathrm{d}r =&
 \frac{1}{k^4\sigma^2} \big[
 6\sin(k\sigma)+(k^2\sigma^2(1-4\lambda+3\lambda^2)
 -6)\sin(k\lambda\sigma) \notag \\
 &-2k\sigma\cos(k\sigma) - k\sigma(4 +
 \lambda(k^2\sigma^2(\lambda-1)^2 - 6))\cos(k\lambda\sigma)\big] \\
  \int_\sigma^{\lambda \sigma} r
 \left( \frac{r}{\sigma}-1\right)^3 \sin (kr) \mathrm{d}r =&
 \frac{1}{k^5\sigma^3} \big[k\sigma(6\sin(k\sigma) + (18-24\lambda
   +k^2\sigma^2(\lambda-1)^2(4\lambda-1))\sin(k\lambda\sigma)) \notag \\
   &+24\cos(k\sigma) + (6k^2\sigma^2(\lambda-1)(2\lambda-1)-\lambda
   k^4\sigma^4(\lambda-1)^3)\cos(k\lambda\sigma)\big] \\
  \int_\sigma^{\lambda \sigma} r
 \left( \frac{r}{\sigma}-1\right)^4 \sin (kr) \mathrm{d}r =&
 \frac{1}{k^6\sigma^4}
 \big[(k^2\sigma^2(\lambda-1)(36-60\lambda+k^2\sigma^2(\lambda - 1)^2(5\lambda-1))
   + 120)\sin(k\lambda\sigma) - 120\sin(k\sigma) \notag \\
   &+k\sigma(24\cos(k\sigma) -
   (24(5\lambda-4)-4k^2\sigma^2(\lambda-1)^2 (5\lambda-2) +
   \lambda k^4\sigma^4(\lambda-1)^4))\cos(k\lambda\sigma) \big]
\end{align}
\end{widetext}
\subsubsection{Averaged pair distribution  $g_\sigma(\rr)$}

@@

We begin with the definition of $A_1$,
\begin{align}
  A_1 = \frac{1}{2}\iint n^{(2)}(\rr_1,\rr_2) \Phi(|\rr_1-\rr_2|)
  \dd\rr_1 \dd\rr_2
\end{align}

But we must use the relation
\begin{align}
  n^{(2)}(\rr_1,\rr_2) = n(\rr_1)n(\rr_2)g^{(2)}(\rr_1,\rr_2)
\end{align}
and equation \ref{eq:g2-our-mean} so that
\begin{align}
  A_1 &=\frac{1}{4}\iint n(\rr_1)n(\rr_2)\Big[
    g_s(r_{12},g_\sigma(\rr_1)) \notag \\
    &\hspace{2cm} + g_s(r_{12},g_\sigma(\rr_2)) \Big] \Phi(r_{12})
  \dd\rr_1 \dd\rr_2.
\end{align}
$g_s$ is symmetric in the exchange of $\rr_1$ and $\rr_2$, and we'll
expand $g_s$ into its polynomial form
\begin{align}
  A_1 &= \frac{1}{2}\iint n(\rr_1)n(\rr_2)g_s(r_{12},g_\sigma(\rr_1))
  \Phi(r_{12}) \dd\rr_1 \dd\rr_2 \\
  &= \frac{1}{2}\iint n(\rr_1)n(\rr_2)\sum_i
  \gamma_i(g_\sigma(\rr_1))\xi_i(r_{12})\Phi(r_{12}) \dd\rr_1 \dd\rr_2 \\
  &= \frac{1}{2} \sum_i \iint \dd\rr_1 \dd\rr_2
  n(\rr_1)n(\rr_2)\gamma_i(g_\sigma(\rr_1)) \xi_i(r_{12})\Phi(r_{12}).
\end{align}
Then we can rearrange things a little better
\begin{align}
  A_1 &= \frac{1}{2} \sum_i \int \dd\rr_1 n(\rr_1)
  \gamma_i(g_\sigma(\rr_1)) \int \dd\rr_2 n(\rr_2)
  \xi_i(r_{12}) \Phi(r_{12})
\end{align}
\fixme{We have introduced an approximation to the pair distribution
  function above... here we discuss an implementation of thermodynamic
  perturbation theory for an inhomogeneous system using this pair
  distribution function.}
